---
title: "Text Mining"
author: "Kavya Gautam"
date: "December 5, 2015"
output: html_document
---
Text Mining. The R package tm provides functionality for text mining applications. The data directories “rec.autos” and “rec.motorcycles” contain a subset of the 20 Newsgroups data set. 

```{r}

library(tm)
?Corpus
?DirSource

```

list.files("/Users/kavyagautam/R Prog/281HW4/Newsgroup_data", all.files = T, recursive = T)
news.corpus <- Corpus(DirSource("/Users/kavyagautam/R Prog/281HW4/Newsgroup_data",recursive=T), readerControl = list(reader = readPlain))


(a) Use the DirSource() and Corpus() tm commands to load the data into R (you may want to use the reader = readPlain argument in the call to Corpus()). Confirm all the documents were read into the corpus by 


```{r}
news.corpus <- Corpus(DirSource("/Users/kavyagautam/R Prog/281HW4/Newsgroup_data",recursive=T), readerControl = list(reader = readPlain))

#Showing the output of length(news.corpus)
length(news.corpus)

#Printing the corpus entry corresponding to document rec.autos/103806 (hint: use names(news.corpus) and [[]] indexing to access corpus entries). 
names(news.corpus)
news.corpus[["103806"]]
writeLines(as.character(news.corpus[["103806"]]))

#Storing the names of news.corpus for later use
filename.index <- names(news.corpus)
```


(b) Use the tm_map() command to apply the following preprocessing transformations of the corpus (in order): removePunctuation, removeNumbers, tolower, removeWords (with argument stopwords("english")) and stripWhitespace. After each step, print document rec.autos/103806 to confirm the effect of the operation

For the next step "tolower" - is not part of the getTransformations() list, since "tm_map" doesnt operate on simple character values anymore in new R version. so we need to use the "content_transformer" function wrapper to convert everything to the correct data type within the corpus


```{r}
#TRANSFORMATIONS OF THE CORPUS IN ORDER - Printing rec.autos/103806 to confirm the effect of the operation after eachs step

#removePunctuation
news.corpus <- tm_map(news.corpus, removePunctuation)
as.character(news.corpus[["103806"]])

#removeNumbers
news.corpus <- tm_map(news.corpus, removeNumbers)
as.character(news.corpus[["103806"]])

#tolower
news.corpus <- tm_map(news.corpus, content_transformer(tolower))
as.character(news.corpus[["103806"]])

#removeWords (with argument stopwords("english"))
news.corpus <- tm_map(news.corpus, removeWords, stopwords("english"))
as.character(news.corpus[["103806"]])

# And stripWhitespace
news.corpus <- tm_map(news.corpus, stripWhitespace)
as.character(news.corpus[["103806"]])

```

(b) continued....
After the last transformation, please use the command: news.corpus <- Corpus(VectorSource(news.corpus)) to make sure the corpus elements are left in plain text prior to the next step.


```{r}
#Finally after all the transformations, to make sure the corpus elements are left in plain text prior to the next step.
news.corpus <- Corpus(VectorSource(news.corpus))

```

(c) Use the DocumentTermMatrix() command to create a document-term matrix with TFIDF weights (also use minWordLength = 1 and minDocFreq = 1 arguments).  

docTermMatrix = DocumentTermMatrix(news.corpus, control = list(weighting = weightTfIdf,minWordLength = 1,minDocFreq = 1))

```{r}
docterm.mat <- DocumentTermMatrix(news.corpus, control=list(weighting = function(x) weightTfIdf(x, normalize = FALSE), minWordLength = 1, minDocFreq = 1))
docterm.mat
inspect(docterm.mat[1:5,1:20])
```
all.equal(docterm.mat, docTermMatrix)

(C) contd...., Then
 
```{r}
#•Use dim() to report the dimensions of the resulting matrix. This is a very sparse matrix with far fewer non-zero entries than the dim() values would suggest.
dim(docterm.mat)

#•Use the inspect() command to confirm that the terms 'bmw' and 'clutch' are present in our rec.autos/103806 document but 'mother' isn’t.
#--------------------------------------------------------------------------
# In order to access the record corresponding to our rec.autos/103806 document, we need the index of the record "103806". we can fetch that using our filename.index vector we stored earlier as follows:
which(filename.index == 103806)

#Lets inspect the record now using the above index - 980 (which corresponds to the rec.autos/103806 document)
inspect(docterm.mat[980, c("bmw", "clutch", "mother")])

```